---
title: "Collaborative Project"
subtitle: "User Testing - Survey Analysis"
author: "Team Name / Student Name"
date: "`r format(Sys.Date(), '%B %d %Y')`"
toc: true
execute:
  echo: false
  warning: false
---

```{r setup, include=FALSE}
# Load libraries
library(tidyverse)
library(psych)    # For descriptive stats
library(ggpubr)   # For enhanced plots like corr plots
```

## 1. Load and Inspect Data

```{r loading, include=FALSE}
# Load CSV (pre-generated sample data)
survey_data <- read_csv("data/2023_24_survey_responses.csv")

# View structure and preview
glimpse(survey_data)
```

```{r somedata}
head(survey_data)
```

## 2. Descriptive Statistics

### Summary for all numeric columns

```{r}
# Summary for all numeric columns
describe(survey_data)
```

### Summary for a specific column (e.g. enjoyment)

```{r}
# Summary for a specific column (e.g. enjoyment)
summary(survey_data$enjoyment)
```

## 3. Test for Normality

```{r}
# Histogram and Shapiro-Wilk test for 'enjoyment'
ggplot(survey_data, aes(x = enjoyment)) + 
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Enjoyment Ratings")

shapiro.test(survey_data$enjoyment)
```

## 4. Boxplots and Comparisons

```{r}
# Boxplot of enjoyment by gender
ggplot(survey_data, aes(x = gender, y = enjoyment)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Enjoyment by Gender")
```

## 5. Correlation Analysis

```{r}
# Scatterplot and correlation test between immersion and engagement
ggscatter(survey_data, x = "immersion", y = "engagement", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson") 

cor.test(survey_data$immersion, survey_data$engagement)
```

## 6. Observations and Next Steps

-   The **enjoyment** scores appear to be normally distributed / not normally distributed.
-   There is / isn't a significant correlation between immersion and engagement.
-   Suggestions for game improvement based on results...
