---
title: "Performing Survey Analysis"
subtitle: "Survey Data Analysis (2022–23)"
author: "Niall McGuinness"
date: "04/10/2025"
---

```{r setup, include=FALSE}
# Load tidyverse for wrangling and visualization
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)  # For data wrangling, plotting — includes ggplot2, dplyr, readr, etc.

# Load survey data
survey_data <- read.csv("data/2022_23_streaming_survey_responses.csv", stringsAsFactors = FALSE)
```

# Introduction

This report presents a descriptive analysis of data collected from a survey on streaming habits. Participants answered questions about their viewing behavior, mood, and platform preferences.

We use basic statistics and visualizations to explore relationships between variables and identify patterns.

---

# 1. Dataset Overview

We begin by exploring the structure of the dataset.

## 1.1 Preview of the Data

```{r}
head(survey_data)
```

## 1.2 Dataset Structure

```{r}
str(survey_data)
```

---

# 2. Descriptive Statistics

Here, we summarize numerical variables using measures like mean, median, and standard deviation.

## 2.1 Summary of Key Variables

```{r}
summary(select(survey_data, Age, Subscriptions, Hours.per.Week, 
               Last.Binge.Session..episodes.movies.,
               Would.Recommend..1.10., Mood.Matching.Importance))
```

## 2.2 Central Tendency and Spread

```{r}
mean_age <- mean(survey_data$Age, na.rm = TRUE)
median_hours <- median(survey_data$Hours.per.Week, na.rm = TRUE)
sd_hours <- sd(survey_data$Hours.per.Week, na.rm = TRUE)

paste("Mean Age:", mean_age)
paste("Median Streaming Hours per Week:", median_hours)
paste("Standard Deviation of Hours per Week:", round(sd_hours, 2))
```

---

# 3. Frequency Tables

These tables summarize categorical responses.

## 3.1 Preferred Streaming Platforms

```{r}
table(survey_data$Main.Platform)
```

## 3.2 Binge-Watching Habits

```{r}
table(survey_data$Binged.Over.5)
```

---

# 4. Correlation Between Mood-Related Items

We examine relationships between Likert-scale items (1–5 or 1–10 ratings).

## 4.1 Correlation Matrix

```{r}
likert_data <- select(survey_data,
                      Recommends.Good.Shows,
                      Interface.Usability,
                      Feels.Overwhelmed,
                      Helps.Me.Relax)

cor_matrix <- cor(likert_data, use = "complete.obs")
round(cor_matrix, 2)
```

---

# 5. Data Visualizations

## 5.1 Histogram: Weekly Streaming Hours

```{r}
hist(survey_data$Hours.per.Week,
     breaks = 10,
     col = 'steelblue',
     main = 'Weekly Streaming Hours',
     xlab = 'Hours per Week',
     ylab = 'Frequency')
```

## 5.2 Boxplot: Age by Platform

```{r}
boxplot(Age ~ Main.Platform,
        data = survey_data,
        col = 'orange',
        main = 'Age Distribution by Platform',
        ylab = 'Age')
```

## 5.3 Scatterplot: Mood Matching vs Relaxation

```{r}
plot(survey_data$Mood.Matching.Importance,
     survey_data$Helps.Me.Relax,
     pch = 19,
     col = 'darkgreen',
     main = 'Mood Matching vs Relaxation',
     xlab = 'Mood Matching (1–10)',
     ylab = 'Relaxation Rating (1–5)')
grid()
```

---

# 6. Categorical Visuals

## 6.1 Bar Plot: Platform Usage

```{r}
barplot(table(survey_data$Main.Platform),
        col = 'skyblue',
        main = 'Most Used Streaming Platforms',
        ylab = 'Number of Users',
        las = 2)
```

## 6.2 Pie Chart: Binge Watching Behavior

```{r}
pie(table(survey_data$Binged.Over.5),
    col = c('lightgreen', 'tomato'),
    main = 'Proportion of Binge (>5 Episodes) Viewers')
```

---

# 7. Mood and Binge Behavior

We compare mood-matching importance across binge-watching behavior.

```{r}
boxplot(Mood.Matching.Importance ~ Binged.Over.5,
        data = survey_data,
        col = c('violet', 'lightcoral'),
        main = 'Mood Matching by Binge Habits',
        xlab = 'Binged Over 5?',
        ylab = 'Mood Matching (1–10)')
```

---

# 8. Enhanced Strip Chart

This strip chart provides a visual overview of individual responses for streaming hours.

```{r}
stripchart(survey_data$Hours.per.Week,
           method = "jitter",
           pch = 21,
           col = "blue",
           bg = "lightblue",
           jitter = 0.2,
           main = "Streaming Hours per Week (Individual Respondents)",
           xlab = "Hours Watched per Week",
           xlim = c(0, max(survey_data$Hours.per.Week, na.rm = TRUE)))
abline(h = 1, col = "darkgray", lty = "dashed")
mtext("Each point represents one survey response. Jitter prevents overlap.", side = 3, line = 0.5, cex = 0.8)
```

---

# Conclusion

This report offers a foundation for understanding trends in streaming behaviors. Future work could include hypothesis testing, comparisons across groups, or predictive models for user engagement.
